### 2023-12-26 documentation/thesis/thesis_document.tex
more thesis

@@

The main aim of this project is to develop a low-cost Hall sensor
interface that is also universally expandable. The focus is on mapping
different sensors and being compatible with different magnet types and
shapes. This ensures broad applicability in different scenarios. Another
goal is reproducibility to ensure consistent results. Easy communication
with standard PC hardware maximises the user-friendliness of the
interface. The flexibility to support different sensors and magnets
makes the system versatile and opens up possibilities for different
applications. This low-cost Hall sensor interface will therefore not
only be economically attractive, but also facilitate the integration of
Hall sensors in different contexts. In addition, the low-cost Hall
sensor interface will serve as a development platform for the data
evaluation \gls{mrp} library and provide real measurement data from
magnets. In addition, the interface firmware creates a basis for the
development of a data protocol for exchanging measured values. This
makes it easy to integrate your own measuring devices into the \gls{mrp}
ecosystem at a later date. This is only possible with a minimal
functional hardware and firmware setup and was developed for this
purpose first.

@@

A decisive step in my work was the in-house PCB
design\ref{1D_sensor_schematic_and_circuit_board.png}, which leads out
all the important connections. The placement and alignment of the
components optimised the wiring, which in turn improved the reliability
of the sensor tests. Particular attention was paid to the implementation
of a \passthrough{\lstinline!SYNC!}-\gls{gpio}, which enables subsequent
multi-sensor synchronisation. This functionality opens up the
possibility of synchronising data from different sensors to achieve
precise and coherent measurement results. Overall, this integrated
approach represents an effective solution for the flexible evaluation of
sensors and helps to optimise the development process.

@@

The firmware is responsible for detecting the possible connected
sensors\ref{Implemented_digital_halleffect_sensors.csv} and reading them
out. This measured data can be forwarded to a host \gls{pc} via a user
interface and can then be further processed there. An important
component is that as many common sensors as possible can be easily
connected without having to adapt the firmware. This modularity was
implemented using \passthrough{\lstinline!abstract!} class design. These
are initiated according to the sensors found at startup. If new hardware
is to be integrated, only the required
functions\ref{lst:CustomSensorClass} need to be implemented.

@@

Now that the setup process is complete, the system switches to an
infinite loop, which processes several possible actions. One task is, to
react to user commands which can be sent to the system by the user via
the integrated \gls{cli}. All sensors are read out via a timer interval
set in the setup procedure and their values are stored in a ringbuffer.
Ring buffers, offer efficient data management in limited memory. Its
cyclic structure enables continuous overwriting of older data, saves
memory space and facilitates seamless processing of real-time data. Ring
buffers are well suited for applications with variable data rates and
minimise the need for complex memory management. The buffer can be read
out by command and the result of the measurement is sent to the host.
Each sensor measurement result is transmitted from the buffer to the
host together with a time stamp and a sequential number. This ensures
that in a multi-sensor setup with several sensors. The measurements are
synchronized\ref{sensor-syncronsisation-interface} in time and are not
out of sequence or drift.

@@

\begin{itemize}
\tightlist
\item
  grafik zeigen
\item
  einzelne module erleutern
\end{itemize}

\hypertarget{hal}{%
\subsubsection{HAL}\label{hal}}

@@

\hypertarget{visualisation}{%
\subsubsection{Visualisation}\label{visualisation}}

@@

\hypertarget{meta-data}{%
\subsubsection{Meta-Data}\label{meta-data}}

\hypertarget{multible-sensor-setup}{%
\subsection{Multible sensor setup}\label{multible-sensor-setup}}

@@

\hypertarget{network-proxy}{%
\subsubsection{Network-Proxy}\label{network-proxy}}

@@

\hypertarget{sensor-syncronisation}{%
\subsubsection{Sensor Syncronisation}\label{sensor-syncronisation}}

@@

\hypertarget{command-router}{%
\subsubsection{Command-Router}\label{command-router}}

@@

\hypertarget{examples}{%
\subsection{Examples}\label{examples}}

@@

\begin{lstlisting}[language=Python, caption={User implemented custom find most similar readings algorithm}, label=lst:custom_find_similar_values_algorithm]
    @staticmethod
    def customFindSimilarValuesAlgorithm(_readings: [MRPReading.MRPReading], IP_return_count: int = -1) -> [MRPReading.MRPReading]:
        import heapq
        heap = []
        # SET RESULT VALUE COUNT
        IP_return_count = max([int(IP_return_count),len(_readings)])
        if IP_return_count < 0:
            IP_return_count = len(_readings) / 5
        # CALCULATE TARGET VALUE: MEAN FROM ALL VALUES
        target_value = 0.0
        for idx, r in enumerate(_readings):
            mean: float = MRPAnalysis.MRPAnalysis.calculate_mean(r)
            target_value = target_value + mean
        target_value = target_value / len(_readings)
        # PUSH READINGS TO HEAP
        for value in _readings:
            # USE DIFF AS PRIORITY VALUE IN MIN-HEAP
            diff = abs(value - target_value)
            heapq.heappush(heap, (diff, value))
        # RETURN X BEST ITEMS FROM HEAP
        similar_values = [item[1] for item in heapq.nsmallest(IP_return_count, heap)]
        # CLEAN UP USED LIBRARIES AND RETURN RESULT
        del heapq
        return similar_values
\end{lstlisting}

@@

Once the filter function has been implemented, it still needs to be
integrated into the analysis
pipeline\ref{lst:pipeline_mrp_evaluation_yaml}. Here, the example
pipeline \ref{example_measurement_analysis_pipeline.png} is simplified
and an additional stage \passthrough{\lstinline!find\_similar\_values!}
has been added, which has set
\passthrough{\lstinline!customFindSimilarValuesAlgorithm!} as the
function to be called. As a final step, the result is used in the
\passthrough{\lstinline!plot\_filtered!} stage for visualisation.

@@

\hypertarget{problems}{%
\section{Problems}\label{problems}}

@@

\begin{itemize}
\tightlist
\item
  magfield camera
\end{itemize}

>>>>

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  State of the art
\item
  Usecases
\item
  Unified Sensor
\item
  Software readout framework
\item
  Usability improvements
\item
  Evaluation
\end{enumerate}

@@

The main aim of this project is to develop a low-cost Hall sensor
interface that is also universally expandable. The focus is on mapping
different sensors and being compatible with different magnet types and
shapes. This ensures broad applicability in different scenarios. Another
goal is reproducibility to ensure consistent results. Easy communication
with standard PC hardware maximises the user-friendliness of the
interface. The flexibility to support different sensors and magnets
makes the system versatile and opens up possibilities for different
applications. A low-cost Hall sensor interface will therefore not only
be economically attractive, but also facilitate the integration of Hall
sensors in different contexts. In addition, the low-cost Hall sensor
interface will serve as a development platform for the data evaluation
\gls{mrp} library and provide real measurement data from magnets. In
addition, the interface firmware creates a basis for the development of
a data protocol for exchanging measured values. This makes it easy to
integrate your own measuring devices into the \gls{mrp} ecosystem at a
later date. This is only possible with a minimal functional hardware and
firmware setup and was developed for this purpose first.

@@

A decisive step in my work was the in-house PCB
design\ref{1D_sensor_schematic_and_circuit_board.png}, which leads out
all the important connections. The placement and alignment of components
optimised the wiring, which in turn improved the reliability of the
sensor tests. Particular attention was paid to the implementation of a
\passthrough{\lstinline!SYNC!}-\gls{gpio}, which enables subsequent
multi-sensor synchronisation. This functionality opens up the
possibility of synchronising data from different sensors to achieve
precise and coherent measurement results. Overall, this integrated
approach represents an effective solution for the flexible evaluation of
sensors and helps to optimise the development process.

@@

The firmware is responsible for detecting the possible connected
sensors\ref{Implemented_digital_halleffect_sensors.csv} and reading them
out. This measured data can be forwarded to a host \gls{pc} via a user
interface and can then be further processed there.

An important component is that as many common sensors as possible can be
easily connected without having to adapt the firmware. This modularity
was implemented using \passthrough{\lstinline!abstract!} class design.
These are initiated according to the sensors found at startup. If new
hardware is to be integrated, only the required
functions\ref{lst:CustomSensorClass} need to be implemented.

@@

Now that the setup process is complete, the system switches to an
infinite loop, which processes several possible actions. One task is, to
react to user commands which can be sent to the system by the user via
the integrated \gls{cli}. All sensors are read out via a timer interval
set in the setup procedure and their values are stored in a ringbuffer.
Ring buffers, offer efficient data management in limited memory. Its
cyclic structure enables continuous overwriting of older data, saves
memory space and facilitates seamless processing of real-time data.

Ring buffers are well suited for applications with variable data rates
and minimise the need for complex memory management. The buffer can be
read out by command and the result of the measurement is sent to the
host. Each sensor measurement result is transmitted from the buffer to
the host together with a time stamp and a sequential number. This
ensures that in a multi-sensor setup with several sensors. The
measurements are synchronized\ref{sensor-syncronsisation-interface} in
time and are not out of sequence or drift.

@@

\hypertarget{extension-modules}{%
\section{Extension Modules}\label{extension-modules}}

Im folgenden werden die einzelnen module, welche vom user modifiziert
und ersetzt werden können im details erleutert.

\hypertarget{hal}{%
\subsection{HAL}\label{hal}}

@@

\hypertarget{visualisation}{%
\subsection{Visualisation}\label{visualisation}}

\hypertarget{storage}{%
\subsection{Storage}\label{storage}}

\begin{itemize}
\tightlist
\item
  metadata management
\end{itemize}

@@

\hypertarget{multible-sensor-setup}{%
\section{Multible sensor setup}\label{multible-sensor-setup}}

@@

\hypertarget{network-proxy}{%
\subsection{Network-Proxy}\label{network-proxy}}

@@

\hypertarget{sensor-syncronisation}{%
\subsection{Sensor Syncronisation}\label{sensor-syncronisation}}

@@

\hypertarget{command-router}{%
\subsection{Command-Router}\label{command-router}}

@@

\hypertarget{examples}{%
\section{Examples}\label{examples}}

The following shows some examples of how the \gls{mrp} library can be
used. These examples are limited to a functional minimum for selected
modules of the \gls{mrp} library. The documentation\ref{documentation}
contains further and more detailed examples. Many basic examples are
also supplied in the form of test cases\ref{tests}.

\hypertarget{mrpanalysis}{%
\subsection{MRPAnalysis}\label{mrpanalysis}}

Dieses beispiel zeigt

\begin{lstlisting}[language=Python]
from MRP import MRPAnalysis
from MRP import MRPDataVisualization
# Create a empty reading with no settings. Only the raw values are needed, no metadata
reading = MRPReading.MRPReading()
# CREATE A SAMPLE MEASUREMENT
for i in range(1000):
    measurement = MRPReadingEntry.MRPReadingEntry()
    # readout sensor or use dummy data and assign result
    measurement.value = (random.random() -0.5) * 2
    reading.insert_reading_instance(measurement, False)

# PLOT MEAN VISUALIZATION
MRPDataVisualization.MRPDataVisualization.plot_error([reading])
# APPLY COMPENSATION
# Here the ``calculate_mean`` function is used
# see MRPAnalysis module for alternatives
reading_mean_value = MRPAnalysis.MRPAnalysis.calculate_mean(reading)
# we want to subtract the mean value from all readings
reading_mean_value = -reading_mean_value
# modify measurement values
MRPAnalysis.MRPAnalysis.apply_global_offset_inplace(reading, reading_mean_value)
\end{lstlisting}

\hypertarget{mrphallbacharraygenerator}{%
\subsection{MRPHallbachArrayGenerator}\label{mrphallbacharraygenerator}}

\begin{figure}
\centering
\includegraphics{./generated_images/border_Generated_hallbach_array_with_generated_cutouts_for_eight_magnets.png}
\caption{Generated hallbach array with generated cutouts for eight
magnets
\label{Generated_hallbach_array_with_generated_cutouts_for_eight_magnets.png}}
\end{figure}

The following example code\ref{lst:mrpexample_hallbach}, shows how a
simple Hallbach magnetic ring can be generated. Eight random
measurements are generated here. It is important that the magnet type
(here \passthrough{\lstinline!N45\_CUBIC\_15x15x15!}) is specified. This
is necessary so that the correct magnet cutouts can be generated when
creating the 3D model. After the measurements have been generated, they
are provided with a position and rotation offset according to the
Hallbach design and calucation
scheme\cite{HallbachMagnetDesignPaper} using the
\passthrough{\lstinline!MRPHallbachArrayGenerator!} module.

\begin{lstlisting}[language=Python, caption={MRPHallbachArrayGenerator example for generating an OpenSCAD based hallbach ring}, label=lst:mrpexample_hallbach]
readings = []
for idx in range(8):
  # GENERATE EXAMPLE READINGS USING N45 CUBIC 15x15x15 MAGNETS
  readings.append(MRPSimulation.MRPSimulation.generate_reading(MRPMagnetTypes.MagnetType.N45_CUBIC_15x15x15))
## GENERATE HALLBACH
hallbach_array: MRPHallbachArrayGenerator.MRPHallbachArrayResult = MRPHallbachArrayGenerator.MRPHallbachArrayGenerator.generate_1k_hallbach_using_polarisation_direction(readings)
# EXPORT TO OPENSCAD
## 2D MODE DXF e.g. for lasercutting
MRPHallbachArrayGenerator.MRPHallbachArrayGenerator.generate_openscad_model([hallbach_array], "./2d_test.scad",_2d_object_code=True)
## 3D MODE e.g. for 3D printing
MRPHallbachArrayGenerator.MRPHallbachArrayGenerator.generate_openscad_model([hallbach_array], "./3d_test.scad",_2d_object_code=False)
\end{lstlisting}

In the last step, a 3D model with the dimensions of the magnet type set
is generated from the generated magnet positions. The result is an
\passthrough{\lstinline!OpenSCAD!}\cite{OpenSCAD} file, which
contains the generated \passthrough{\lstinline!OpenSCAD!}
sourcecode\ref{Generated_hallbach_array_with_generated_cutouts_for_eight_magnets.png}.

@@

\begin{lstlisting}[language=Python, caption={User implemented custom find most similar readings algorithm}, label=lst:custom_find_similar_values_algorithm]
    @staticmethod
    def FindSimilarValuesAlgorithm(_readings: [MRPReading.MRPReading], IP_return_count: int = -1) -> [MRPReading.MRPReading]:
        import heapq
        heap = []
        # SET RESULT VALUE COUNT
        IP_return_count = max([int(IP_return_count),len(_readings)])
        if IP_return_count < 0:
            IP_return_count = len(_readings) / 5
        # CALCULATE TARGET VALUE: MEAN FROM ALL VALUES
        target_value = 0.0
        for idx, r in enumerate(_readings):
            mean: float = MRPAnalysis.MRPAnalysis.calculate_mean(r)
            target_value = target_value + mean
        target_value = target_value / len(_readings)
        # PUSH READINGS TO HEAP
        for value in _readings:
            # USE DIFF AS PRIORITY VALUE IN MIN-HEAP
            diff = abs(value - target_value)
            heapq.heappush(heap, (diff, value))
        # RETURN X BEST ITEMS FROM HEAP
        similar_values = [item[1] for item in heapq.nsmallest(IP_return_count, heap)]
        # CLEAN UP USED LIBRARIES AND RETURN RESULT
        del heapq
        return similar_values
\end{lstlisting}

@@

Once the filter function has been implemented, it still needs to be
integrated into the analysis
pipeline\ref{lst:pipeline_mrp_evaluation_yaml}. Here, the example
pipeline \ref{example_measurement_analysis_pipeline.png} is simplified
and an additional stage \passthrough{\lstinline!find\_similar\_values!}
has been added, which has set
\passthrough{\lstinline!FindSimilarValuesAlgorithm!} as the function to
be called. As a final step, the result is used in the
\passthrough{\lstinline!plot\_filtered!} stage for visualisation.

@@

This work describes the development of a universal Python library that
is used to efficiently process data from magnetic field sensors from
acquisition to analysis. In order to ensure a practical application and
to give users the opportunity to directly acquire their own magnetic
field data, cost-effective and easily reproducible hardware was also
developed.

The hardware is based on widely used magnetic field sensors and low-cost
microcontrollers, which enables an easily expandable and applicable
solution for measuring magnets with repeatable accuracy. A particular
focus was placed on expandability by the user. Interchangeable modules
allow the user to develop their own analysis algorithms without having
to redesign everything from scratch.

This extensibility and customisability was successfully demonstrated
during the evaluation. This underlines the performance of the developed
framework and shows that it is not only effective in the processing of
magnetic field sensor data, but also offers a flexible platform for the
implementation of user-specific analyses.

@@

In this version of the framework, a solid foundation has been laid that
includes all the necessary features and is ready for use. During
development, particular emphasis was placed on comprehensive
documentation to make it easier to get started. Together with examples
for various use cases, a user can quickly evaluate the framework.

However, it should be noted that the framework is still in its infancy.
The stable version distributed via the package registry is well suited
for the intended purpose. All tests and evaluations took place under
normal conditions, especially for the developed hardware sensors, as the
\gls{mrp} library works successfully with the measurement data.

On the software side, the focus is on integration for the support of
more professional measuring devices. Only in this way is it possible to
evaluate and improve the sensor hardware and quantify the measurement
results.

To summarise, it can be said that a solid software framework has been
created that can be used directly for the intended purpose. It provides
a good foundation, but can be further developed by integrating
professional measurement devices to enable a more comprehensive
evaluation and improvement of the sensor hardware.
